
%\documentclass[12pt,preprint]{aastex}
\documentclass{emulateapj}
\usepackage{url}
%\usepackage{natbib}
%\usepackage{xspace}
\def\arcsec{$^{\prime\prime}$}
\bibliographystyle{apj}
\newcommand\degree{{^\circ}}
\newcommand\surfb{$\mathrm{mag}/\square$\arcsec}
\newcommand\Gyr{\rm{~Gyr}}
\newcommand\msun{\rm{M}_\odot}
\newcommand\kms{km s$^{-1}$}
\newcommand\al{$\alpha$}
\newcommand\ha{$\rm{H}\alpha$}
\newcommand\hb{$\rm{H}\beta$}



\shorttitle{Short Title}
\shortauthors{Yoachim et al.}

\begin{document}

\title{Publication Lifetimes of American Astronomy PhDs}

\author{Peter Yoachim\altaffilmark{1}, author 2, author 3 
  }
\altaffiltext{1}{Department of Astronomy, University of Washington, Box 351580,
Seattle WA, 98195; {yoachim@uw.edu} }




\begin{abstract}
blah blah blah, abstract

\end{abstract}
\keywords{employment}



\section{Introduction}

\citet{Seth09} build on the work of \citet{Metcalfe08} and looks at the rise of the postdoc era in Astronomy and Astrophysics.  They show that as the infaltion-adjusted US-Astronomy budget doubled over 20 years, there was a modest increase in PhD production and a huge rise in the number of temporary postdoc positions.

In this paper, we use the ADS to explore the ramifications of that rising postdoc bubble.  While the NSF regularly surveys PhD reciepents (e.g., \url{http://nsf.gov/statistics/nsf07305/}), even with high participation rates, surveys will tend to be biased against those that leave the field since they will be harder to contact.

Thus, we propose to use the publication record of PhDs to track the overall health of the field.  

Unfortunately, some astronomers still believe that ``... the [astronomy faculty job] market is no worse or better than it is has been for at least a decade or two'' (via Charfman \url{http://jjcharfman.tumblr.com/}).  The goal of this paper is to check this statement and quantify the astronomy job market by looking at the publication footprint PhD Astronomers leave in the ADS system. 

XXX-description of ADS, how it includes peer-reviewed papers as well as things like conference proceedings, posters, grants, etc.

\section{Database Construction}

All the code used for this paper is available on github\footnote{\url{https://github.com/yoachim/AstroHireNetwork}}

We use the excelent Python Module to Interact with NASA's ADS that Doesn't Suck™\footnote{\url{https://github.com/andycasey/ads}} to programatically access records from the SAO/NASA Astrophysics Data System XXX--figure out proper way to cite code. 

Our procedure for building a database of Astronomy and Astrophysics PhDs is as follows.  We select all PhD Thesis records for a given year and screen to include only those records that have an affiliation in the United States.

For each US PhD, we then query ADS for all entries with the PhD author's last name and first initial with a date limit of 7 years pre-PhD to the present (Jan 2016).

For common names (e.g., Jones, R and Williams, B), our author query returns entries from multiple authors.  To try and restrict the results to only those papers with the same author as the PhD thesis, we construct a network graph.

Every ADS entry is a node in our network. We draw an edge between nodes if they meet any of the following criteria:
\begin{itemize}
\item{If the affiliation of the thesis author matches, and the year of the entry is within 2 years. XXX-describe affil match fuzzy string, plus if one string is contained in the other.}
\item{If the entries have 3 or more authors in common in their author lists.}
\item{If there are two authors and they are identical on the two entries.}
\item{If the information contained in the abstracts are similar enough as measured by XXX-magic fucking black boxXXX TfidfVectorizer}
  \item{If the information contained in the titles of the entries are similar enough.}
\end{itemize}

We then select only the entries which are somehow connected to the PhD thesis of interest.  Some example ADS entry network graphs are shown in Figure~XXX.

Once we have selected only those entries we believe are connected to the correct thesis author, we demand that at least one of the entries be in a peer-reviewed astronomy journal (XXX-list of journals) to eliminate physics thesis entries.

Thus, our definition of a US PhD astronomer is effectively a person who recieved a PhD from an institution in the United States and has published at least one paper in an astronomy journal.

We record the thesis author's most recent ADS entry date (need not be a peer-reviewed paper)  and a bunch of other stuff.

We also flag if there are any other thesis entries (from any year) that match the last name, first initial of the thesis entry so that we can compare the full database to those with more unique names who should suffer fewer errors in the entry network construction.

One known issue is that ADS does not contain all Astronomy PhD dissertations. We found that XXX is not listed in ADS (see github issue).


Plots to make:
\begin{itemize}
\item{Example paper clouds}
\item{number of US astro phds per year}
\end{itemize}




Possible ways we could fail to recognize an author has published for as long as they actually have
\begin{itemize}
\item{If someone simultaneously changes institution and area of study}
  \item{If an astronomer leaves the field for an extended time and re-joins in a different field}
\item{If an astronomer changes their name (although authors can notify ADS of name changes)}
  \item{}
\end{itemize}

Ways we could mistakenly claim an author has published for longer than they actually have
\begin{itemize}
\item{If two authors have similar names and overlap at the same institution}
\item{If two authors have similar names and similar reference lists in their papers}
  \item{If two authors have similar names and similarly named co-authors}
\end{itemize}

The overall point is that errors can happen both ways and both should be relavively rare, giving us a more accurate measurement of astronomer career lifetimes than survey results which rarely top XXX\% response rates.

It is worth emphasizing that while there are undoubtably some cases where we have not accuratly recorded an astronomer, we do not suffer the selection biases that go with survey-based studies and that our errors should be similar accross PhD cohorts, thus we can draw strong conclusions about the relative differences between PhD classes.

\section{Observations}

\section{Discussion}

{\bf{Astronomy is not suffering from a funding crisis, but rather a funding allocation crisis.}} A simple thought experiment can help illustrate the problem. What would happen if the NSF was able to double it's Astronomy budget? Much of that money would be distributed as grants, that would then be used to hire graduate students and post docs. Very little of the money would go to funding permanent positions. Increasing federal astronomy funding would actually exaserbate PhD overproduction. 





%\begin{figure}
%\epsscale{.5}
%\caption{ \emph{continued}.}
%\end{figure}


\section{Conculsions}
blah blah conclusions.


%--------------------------------------
\acknowledgments
Thanks everyone. Thanks to sponsors of AAS Hack Day 2016. 


\bibliography{big_jabref.bib}




\end{document}
